### 2025-03-22 workflow/notebooks/draft.Rmd
added figs3 in draft

---
title: "Structural analysis"
# knit: (function(input_file, encoding) {
#   rmarkdown::render(input_file,
date: "`r format(Sys.time(), '%d-%b-%Y')`"
author: "Giacomo Mutti"
output:
    rmdformats::html_clean:
        number_sections: false
        code_folding: hide
        self_contained: true
        fig_caption: true
        # thumbnails: false
        # lightbox: false
        gallery: true
        toc_depth: 3
        highlight: kate
---

```{r setup, include=FALSE, dev="CairoPNG"}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, cache.lazy = FALSE,
                      crop = TRUE, dev="CairoPNG")
knitr::knit_hooks$set(crop = knitr::hook_pdfcrop)
knitr::opts_knit$set(root.dir = "../../")
```

```{r libraries, message=FALSE, warning=FALSE, cache=FALSE}
library(tidyverse)
library(patchwork)
library(ggtree)
library(cowplot)
library(glue)
library(ggh4x)
library(ggpubr)
library(rstatix)
source("workflow/scripts/functions.R")
theme_set(theme_classic(base_family = "Helvetica"))

params <- yaml::read_yaml("config/params.yaml")
params_dataset <- yaml::read_yaml("config/Hsap_draft.yaml")

homodir <- glue(params$outdir, "homology/", params_dataset$homology_dataset)
phylodir <- glue(params$outdir, "phylogeny/", params_dataset$phylo_dataset)
```

```{r read_stuff, cache=FALSE}
seeds <- readLines(params_dataset$test_seeds)

tax <- read_delim(glue("results/data/meta/",
                       params_dataset$homology_dataset, "_uniprot_genomes.tsv"), 
                  show_col_types=F)
taxmap <- read_delim(glue(homodir, "/db/taxidmap"), 
                     delim = "\t", col_names = c("label", "Tax_ID"),
                     show_col_types = FALSE) %>% 
  left_join(tax, by = "Tax_ID") %>% 
  dplyr::select(label, Proteome_ID, mnemo)

taxidmap <- read_delim(glue(homodir,"/db/taxidmap"), 
                       col_names = c("target", "Tax_ID"), show_col_types=F)

sptree <- read.tree(params_dataset$species_tree)
translate <- deframe(distinct(dplyr::select(taxmap, Proteome_ID, mnemo)))
# sptree$tip.label <- names(translate)[match(translate, sptree$tip.label)]

species_dist_df <- tibble()
for (sp in sptree$tip.label){
  if (sp=="Hsap") {next}
  dist <- castor::get_all_distances_to_tip(sptree, "Hsap")[MRCA(sptree, "Hsap", sp)]
  mrca <- c(sptree$tip.label, sptree$node.label)[MRCA(sptree, "Hsap", sp)]
  species_dist_df <- bind_rows(species_dist_df, tibble(mnemo=sp, value=dist, mrca=mrca))
}

table_columns <- c('Proteome_ID', 'Tax_ID', 'count1', 'count2', 'count3', 
                   'genome', 'source', 'species', 'mnemo')

lddt_df <- read_delim(paste0(params$data_dir, "structures/", 
                             tax$Proteome_ID, "/", tax$Proteome_ID, "_mean_plddt.tsv"),
           col_names = c("target", "mean_lddt"), show_col_types = FALSE)

aln_seeds <- readLines(glue(phylodir,"/ids/", params_dataset$seed, "_aln.ids"))
```

# Homology

```{r read_self_blast}
blast_self <- read_delim(glue(homodir, "/allvall/",
                              params_dataset$seed, "_", params_dataset$seed, "_blast.tsv"),
                         show_col_types=FALSE,
                         col_names = blast_columns) %>% 
  # filter(query %in% seeds) %>%
  filter(query==target) %>% 
  mutate(blast_selfbit=bitscore) %>% 
  select(query, blast_selfbit)

fs_self <- read_delim(glue(homodir, "/allvall/", params_dataset$seed, 
                           "_", params_dataset$seed, "_fs.tsv"), 
                      show_col_types = FALSE,
                         col_names = blast_columns) %>% 
  # filter(query %in% seeds) %>%
  filter(query==target) %>% 
  mutate(fs_selfbit=bitscore) %>% 
  select(query, fs_selfbit)

self <- full_join(blast_self, fs_self, by="query")
```

```{r read_homo, eval=FALSE}
fs <- read_delim(glue(homodir, "/", params_dataset$seed, "_fs.tsv"), 
                 col_names = fs_columns, show_col_types = FALSE) %>% 
  # filter(query %in% seeds) %>%
  mutate(evalue = ifelse(evalue==0, 1e-180, evalue)) %>%
  left_join(taxidmap, by = "target")

blast <- read_delim(glue(homodir, "/", params_dataset$seed, "_blast.tsv"), 
                    col_names = blast_columns, show_col_types = FALSE) %>% 
  # filter(query %in% seeds) %>%
  mutate(evalue = ifelse(evalue==0, 1e-180, evalue))
```

```{r merge_homo, eval=FALSE}
df <- full_join(blast, fs,
                by = c("query", "target"), suffix=c("_Blast", "_Foldseek")) %>%
    filter(query!=target) %>% 
  left_join(self, by="query") %>%
  mutate(pident_Foldseek = 100*pident_Foldseek, 
         qcov_Foldseek = 100*qcov_Foldseek,
         singleton = case_when(is.na(evalue_Foldseek) ~ "only Blast",
                               is.na(evalue_Blast) ~ "only Foldseek", 
                               .default = "Common"))

blast_filt <- filter(blast, 
                         length/qlen*100>params$coverage, 
                         length/slen*100>params$coverage,
                         evalue<as.numeric(params$eval_both)) %>% 
  group_by(query) %>% 
  slice_head(n = params$max_seqs) %>% 
  ungroup()

fs_filt <- filter(fs, 
                      qcov*100>params$coverage, 
                      tcov*100>params$coverage, 
                      evalue<as.numeric(params$eval_both)) %>% 
  group_by(query) %>% 
  slice_head(n = params$max_seqs) %>% 
  ungroup()

df_filt <- full_join(blast_filt, fs_filt,
                by = c("query", "target"), suffix=c("_Blast", "_Foldseek")) %>%
    filter(query!=target) %>% 
  left_join(self, by="query") %>%
  mutate(pident_Foldseek = 100*pident_Foldseek, 
         qcov_Foldseek = 100*qcov_Foldseek,
         singleton = case_when(is.na(evalue_Foldseek) ~ "only Blast",
                               is.na(evalue_Blast) ~ "only Foldseek", 
                               .default = "Common"))

df %>% 
  group_by(singleton) %>% 
  summarise(n(), n()/nrow(df))

df_filt %>% 
  group_by(singleton) %>% 
  summarise(n(), n()/nrow(df_filt))
```

```{r cath, eval=FALSE}
CATH_df <- read_delim(paste0(params$data_dir, "cath/", tax$Proteome_ID, "_cath.tsv"),
                      delim = "\t", id = "proteome", show_col_types = FALSE) %>%
  mutate(proteome = gsub("_cath.tsv", "", basename(proteome))) %>%
         # n_3d_domains = str_count(Gene3D, ";"),
         # n_Pfam_domains = str_count(Pfam, ";")
  mutate(target=paste0("AF-", Entry, "-F1")) %>% 
  rename("CATH"="Gene3D") %>% 
  select(target, Pfam, CATH, -proteome, -Entry) %>% 
  mutate(CAT = gsub("(\\d+\\.\\d+\\.\\d+)\\.\\d+", "\\1", CATH),
         CA = gsub("(\\d+\\.\\d+)\\.\\d+", "\\1", CAT),
         C = gsub("\\.\\d+", "\\1", CA))

threed_df <- read_delim(paste0(params$data_dir, "ids/", tax$Proteome_ID, "_3ds.tsv"),
           delim = "\t", id = "proteome", show_col_types = FALSE) %>% 
  mutate(proteome = gsub("_ids.tsv", "", basename(proteome))) %>% 
  mutate(target=paste0("AF-", Entry, "-F1")) %>% 
  select(-proteome, -Entry) %>% 
  left_join(taxmap, by=c("target"="label")) %>% 
  mutate(sp = gsub("..*_", "", `Entry Name`))

ortho_dbs <- read_delim(paste0(params$data_dir, "ids/", tax$Proteome_ID, "_ids.tsv"),
                      delim = "\t", id = "proteome", show_col_types = FALSE) %>% 
  mutate(proteome = gsub("_ids.tsv", "", basename(proteome))) %>% 
  mutate(target=paste0("AF-", Entry, "-F1")) %>% 
  select(-proteome, -Entry)

ortho_dbs <- full_join(CATH_df, ortho_dbs)
```

## Fig 1

```{r fig1, eval=FALSE}
fig1a <- df %>%
  group_by(singleton) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  mutate(prop=n/sum(n), Filtering="Before") %>%
  bind_rows(df_filt %>%
            group_by(singleton) %>%
            summarise(n = n()) %>%
            ungroup() %>%
            mutate(prop=n/sum(n), Filtering="After")) %>%
  mutate(Filtering=factor(Filtering, labels = c("After", "Before"), ordered = T),
         singleton=case_when(singleton=="only Blast"~"Only Bp", 
                             singleton=="only Foldseek" ~ "Only Fs",
                             .default = singleton)) %>%
  ggplot(aes(singleton, n, fill=singleton)) +
  geom_bar(aes(alpha=Filtering), stat = "identity", position = position_dodge2(reverse=TRUE)) +
  scale_fill_manual(values = palette_singleton_s, guide="none") +
  scale_alpha_manual(values = c(1,.6)) +
  scale_y_continuous(expand = c(0,0), labels = scales::label_number(scale = 1e-6)) +
  labs(fill="Class", x="Class", y="Milions of aligned protein pairs") +
  theme(legend.position = "bottom", legend.box="vertical")

fig1b <- lddt_df %>% 
  ggplot(aes(mean_lddt)) + 
  geom_density(fill=NA, size=1) +
  scale_y_continuous(expand = expansion(mult = c(0, .1))) +
  labs(x="Mean pLDDT", y="") +
  geom_vline(xintercept = 40, linetype=2) +
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.line.y = element_blank())

fig1 <- (fig1a | fig1b) + plot_layout(widths = c(1,2))

ggsave("draft/figures/Fig1_tmp.pdf", fig1, width = 8, height = 4)
```

## Fig 2

```{r fig2, eval=FALSE}
eval_p <- ggplot(df, aes(-log10(evalue_Blast), -log10(evalue_Foldseek))) + 
  stat_bin_2d(breaks=seq(0,20,0.4)) + 
  scale_fill_gradientn(colors=wespal) +
  ggpubr::stat_cor(label.x.npc = .025) +
  coord_cartesian(expand = 0) +
  labs(x="-log10(Evalue BlastP)",y="-log10(Evalue Foldseek)") +
  scale_x_continuous(limits = c(0, 20)) +
  scale_y_continuous(limits = c(0, 20)) +
  theme(legend.position = "none", panel.grid = element_blank()) 
eval_p <- add_marginal(mutate(df, 
                              evalue_Blast=-log10(evalue_Blast), 
                              evalue_Foldseek=-log10(evalue_Foldseek)), 
                       eval_p, "evalue_Blast", "evalue_Foldseek", "singleton", palette_singleton_p)

pident_p <- ggplot(df, aes(pident_Blast, pident_Foldseek)) + 
  stat_bin_2d(breaks=seq(0,100,2)) + 
  scale_fill_gradientn(colors=wespal) +
  ggpubr::stat_cor(label.x.npc = .025) +
  coord_cartesian(expand = 0) +
  scale_x_continuous(limits = c(0, 100)) +
  scale_y_continuous(limits = c(0, 100)) +
  labs(x="% Identity BlastP",y="% Identity Foldseek") +
  theme(legend.position = "none", panel.grid = element_blank())
pident_p <- add_marginal(df, pident_p, "pident_Blast", "pident_Foldseek", "singleton", palette_singleton_p)

qcov_p <- ggplot(df, aes(qcov_Blast, qcov_Foldseek)) + 
  stat_bin_2d(breaks=seq(0,100,2)) + 
  ggpubr::stat_cor(label.x.npc = .025) +
  scale_x_continuous(limits = c(0, 100)) +
  scale_y_continuous(limits = c(0, 100)) +
  labs(x="Query cov. BlastP",y="Query cov. Foldseek") +
  scale_fill_gradientn(colors=wespal) +
  coord_cartesian(expand = 0) +
  theme(legend.position = "none", panel.grid = element_blank()) 
qcov_p <- add_marginal(df, qcov_p, "qcov_Blast", "qcov_Foldseek", "singleton", palette_singleton_p)

lddt_p <- df %>% 
  select(lddt, singleton) %>% 
  ggplot(aes(color=singleton, lddt)) + 
  geom_density() + 
  scale_y_continuous(expand = expansion(c(0.02,0.02))) +
  scale_x_continuous(limits = c(0, 1), 
                     expand = expansion(c(0.02,0.02))) +
  scale_color_manual(values = palette_singleton_p, guide=FALSE) + 
  labs(x="LDDT", color = "Target sets") + 
  theme(legend.position = "none", 
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

alntmscore_p <- df %>% 
  select(alntmscore, singleton) %>% 
  ggplot(aes(color=singleton, alntmscore)) + 
  geom_density() + 
  scale_y_continuous(expand = expansion(c(0.02,0.02))) +
  scale_x_continuous(limits = c(0, 1), 
                     expand = expansion(c(0.02,0.02))) +
  labs(x="TM-Score", color = "Target sets") + 
  scale_color_manual(values = palette_singleton_p, guide=FALSE)+ 
  theme(legend.position = "none", 
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

plddt_p <- df %>% 
  select(target, singleton) %>% 
  mutate(target=gsub("AF-|-F1", "", target)) %>% 
  left_join(lddt_df, by="target") %>% 
  ggplot(aes(color=singleton, mean_lddt)) + 
  geom_density(bw=1) + 
  scale_y_continuous(expand = expansion(c(0.02,0.02))) +
  scale_x_continuous(limits = c(params$low_confidence, 100), 
                     expand = expansion(c(0.02,0.02))) +
  scale_color_manual(values = palette_singleton_p, guide=FALSE) +   
  labs(x="Mean target pLDDT", color = "Target sets") + 
  theme(legend.position = "none", 
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

dbs <- colnames(ortho_dbs)[2:dim(ortho_dbs)[2]]
dbs <- setdiff(dbs, c("InParanoid"))
dbs <- dbs[2:5]

binwidth <- 0.1

jaccard_df <- map_dfr(dbs, ~ {
  col_name <- .x
  select(df_filt, query, target, singleton) %>% 
    left_join(select(ortho_dbs, target, col_name), by=c("query"="target")) %>% 
    left_join(select(ortho_dbs, target, col_name), by=c("target")) %>% 
    rename("query_set"=paste0(col_name, ".x"), "target_set"=paste0(col_name, ".y")) %>% 
    # If any of query or target has na no we dont count
    filter(!is.na(query_set), !is.na(target_set)) %>% 
    mutate(
      jaccard = map2_dbl(strsplit(query_set, ";"), strsplit(target_set, ";"),
                         ~ length(intersect(.x, .y)) / length(union(.x, .y)))
    )  %>% 
    group_by(query, singleton) %>% 
    summarise(avg_Jaccard = mean(jaccard)) %>% 
    mutate(db=col_name)

})

p_dbs <- ggplot(jaccard_df, aes(avg_Jaccard, 1, color=singleton, alpha=db)) +
  stat_ecdf(pad=FALSE) +
  scale_color_manual(values=palette_singleton_p) +
  # scale_linetype_manual(values = c(2,2,2,2,1)) +
  scale_alpha_manual(values = seq(0.3, 1, length.out=4)) +
  coord_fixed() +
  scale_y_continuous(expand = expansion(0.0,0)) +  
  scale_x_continuous(expand = expansion(add = c(0,0.01))) +  
  guides(colour = guide_legend(order = 1),  
         alpha = guide_legend(title = "CATH Level", order = 2)) +
  theme(panel.border = element_rect(fill = NA), 
        legend.position = "bottom",
        panel.grid.major = element_line(linetype = 2)) + 
  # theme_bw(base_family = "Aptos") +
  labs(x="Mean Jaccard Index", y="ECDF", 
       alpha="DB", linetype="DB", color="Target sets")

fig2_first_row <- (pident_p | eval_p | qcov_p)
fig2_second_row <- ((lddt_p / alntmscore_p / plddt_p) | p_dbs) + 
  plot_layout(guides = "collect") & 
  theme(legend.position = 'top')
fig2 <- (fig2_first_row / fig2_second_row) +
  plot_layout(heights = c(1,1.3)) +
  plot_annotation(tag_levels = 'A') &
  theme(plot.tag = element_text(size = 10, face = "bold", family = "Helvetica"))

ggsave("draft/figures/Fig2.pdf", fig2, width = 8, height = 7, device = cairo_pdf)
```

## Fig 3

```{r read_trees}
models_draft <- setdiff(models, c("QT", "FTPY", "FMG", "C10", "C60"))

unrooted_trees <- paste0(params$outdir, 'phylogeny/',
                              params_dataset$phylo_dataset,  "/trees/", 
                              params_dataset$seed, "_unrooted_trees.txt")
unrooted_trees <- "test/flashforward/out/unrooted_trees.txt"

disco_folder <- paste0(params$outdir, 'phylogeny/',
                              params_dataset$phylo_dataset,  "/reco/disco")
disco_folder <- "test/flashforward/reco/"

notung <- paste0(params$outdir, 'phylogeny/',
                              params_dataset$phylo_dataset,  "/reco/", 
                              params_dataset$seed, "_notung.tsv")
notung <- "test/flashforward/out/notung.txt"

apro_folder <- paste0(params$outdir, 'phylogeny/',
                              params_dataset$phylo_dataset,  "/reco/apro")
apro_folder <- "test/flashforward/reco/"

df_trees <- read_delim(unrooted_trees,
                       show_col_types = FALSE,
                       col_names = c("gene", "target", "alphabet", "model", "tree")) %>% 
  mutate(model=factor(model, levels=models)) %>% 
  filter(!is.na(tree)) %>% #, model %in% models_draft) 
  # some FM trees have negative branch lengths
  mutate(tree = gsub(":-", ":", tree)) 
# 722*8*4=23104 trees

# ML trees with selected model and LL
mltrees <- paste0(params$outdir, 'phylogeny/',
                              params_dataset$phylo_dataset,  "/trees/", 
                              params_dataset$seed, "_mltrees.txt")
mltrees <- "test/flashforward/out/mltrees.txt"

df_ml <- read_delim(mltrees,
                    delim = "\t",
                    show_col_types=FALSE) %>%
  # mutate(fn=gsub("_mltrees.txt", "", basename(fn))) %>%
  # separate(fn, c("gene", "target", "alphabet", "model"), sep = "_") %>%
  mutate(model=factor(gsub("3DI", "3Di",
                           gsub("GTR20", "GTR",
                                gsub("\\+.*", "", Model))), levels=models),
         rate=sub("\\+", "", gsub("GTR20|3DI|LG", "", Model)))

disco_fls <- list.files(disco_folder, 
                        pattern = "*output*", full.names = T)
disco_fls <- list.files(disco_folder, 
                        pattern = "^disco", full.names = T)

# Read all trees
ts <- read.tree(text = df_trees$tree)
names(ts) <- paste(df_trees$gene, df_trees$target, df_trees$alphabet, df_trees$model, sep = "_")

disco_rf <- get_disco_rf(disco_fls, sptree)
df_bs <- get_bs_df(ts)

n_tips <- sapply(ts, function(x) length(x$tip.label)) %>% 
  enframe() %>%
  separate(name, c("gene", "target", "alphabet", "model"), sep = "_") %>%
  rename("n_tips"="value")

var_df <- sapply(ts, function(x) { 
                 a <- phytools::midpoint_root(x) 
                 var(ape::dist.nodes(a)[a$edge[1,][1],1:length(a$tip.label)])
                 }) %>% 
  enframe() %>%
  separate(name, c("gene", "target", "alphabet", "model"), sep="_") %>%
  rename("variance_r2t"="value")

# Reconciliation ranger+TCS scores
reco <- read_delim(notung, 
                   show_col_types = FALSE, 
                   col_names = c("gene", "target", "alphabet", "model", "dups", "losses")) %>% 
  left_join(n_tips) %>% 
  left_join(var_df) %>% 
  mutate(model=factor(model, levels=models),
         n_events = (dups+losses)/n_tips) %>% 
  # filter(model %in% models_draft) %>% 
  rename("targets"="target")

apro_files <- list.files(apro_folder, 
                        pattern = "*support.nwk", full.names = TRUE)
apro_files <- list.files(apro_folder, 
                        pattern = "^apro", full.names = TRUE)

apro_trees <- read.tree(text = sapply(apro_files, readLines))
names(apro_trees) <- sapply(str_split(basename(apro_files), "_"), 
                            function(x) paste0(x[2], "_", x[4]))

trees_df <- get_apro_stats(apro_trees)

save.image(file = "workflow/notebooks/eda_cache/until_dftree.Rdata")
```

```{r fig3}
xax <- ifelse(df_model[df_model$model %in% models_draft,]$algorithm=="Distance", "italic", "plain")

DL_df <- reco %>% 
  left_join(df_model) %>% 
  filter(model %in% models_draft) %>% 
  rename("y"="n_events")
DL_test <- stat_test(DL_df, paired = T, alt = "l")

plot_DL <- DL_df %>% 
  ggplot(aes(weave_factors(model, data), y, 
             color=algorithm, fill=targets)) + 
  labs(y="(D+L)/# Tips", x="",fill="Target sets") +
  geom_boxplot(outlier.size = .1) +
  geom_text(aes(label=as.character(p.adj.signif), y=8, x=group2), 
            data = filter(DL_test, p.adj.signif!="ns"), 
            inherit.aes = F, color="grey20") +
  scale_color_manual(values = c("black", "black"),
                     labels = c("ML", expression(italic("Distance")))) +
  guides(x = "axis_nested", 
         colour=guide_legend(title = "Method", override.aes=list(linetype = 0, size=0))) +
  scale_fill_manual(values = palettes_method) +
    theme(
    axis.text.x = element_text(face = xax),
    ggh4x.axis.nestline.x = element_line(color="grey"),
    ggh4x.axis.nesttext.x = element_text(colour = "grey30", 
                                         face = "bold", size=10)
  )

RF_df <- disco_rf %>% 
  mutate(model=factor(model, levels=models)) %>% 
  filter(nsps>length(sptree$tip.label)/2, model %in% models_draft) %>%
  # separate(bn, c("seed", "targets", "alphabet", "model")) %>% 
  left_join(df_model) %>% 
  rename("y"="rf")
RF_test <- stat_test(RF_df, paired = F, alt = "l")

plot_rf <- RF_df %>% 
  ggplot(aes(weave_factors(model, data), y, 
             color=algorithm, fill=targets)) + 
  geom_boxplot(outlier.size = .1) +
  geom_text(aes(label=as.character(p.adj.signif), y=1.05, x=group2), 
            data = filter(RF_test, p.adj.signif!="ns"), 
            inherit.aes = F, color="grey20") +
  labs(x="", y="Norm. RF to species tree", fill="Target sets") +
  scale_color_manual(values = c("black", "black"),
                     labels = c("ML", expression(italic("Distance")))) +
  guides(x = "axis_nested", 
         colour=guide_legend(title = "Method", override.aes=list(linetype = 0, size=0))) +
  scale_fill_manual(values = palettes_method) +
    theme(
    axis.text.x = element_text(face = xax),
    ggh4x.axis.nestline.x = element_line(color="grey"),
    ggh4x.axis.nesttext.x = element_text(colour = "grey30", 
                                         face = "bold", size=10)
  )

apro_df <- trees_df %>% 
  left_join(df_model) %>% 
  select(-y) %>% 
  filter(model %in% models_draft) %>% 
  rename("y"="q1")
apro_test <- stat_test(apro_df, paired = T, alt = "g")

plot_apro <- apro_df %>% 
  ggplot(aes(weave_factors(model, data), y, 
             color=algorithm,fill=targets)) + 
  labs(x="", y="Quartet support", fill="Target sets") + 
  geom_boxplot(outlier.size = .1) +
  geom_text(aes(label=as.character(p.adj.signif), y=1.05, x=group2), 
            data = filter(apro_test, p.adj.signif!="ns"), 
            inherit.aes = F, color="grey20") +
  scale_color_manual(values = c("black", "black"),
                     labels = c("ML", expression(italic("Distance")))) +
  guides(x = "axis_nested", 
         colour=guide_legend(title = "Method", 
                             override.aes=list(linetype = 0, size=0))) +
  scale_fill_manual(values = palettes_method) +
    theme(
    axis.text.x = element_text(face = xax),
    ggh4x.axis.nestline.x = element_line(color="grey"),
    ggh4x.axis.nesttext.x = element_text(colour = "grey30", 
                                         face = "bold", size=10)
  )

fig3 <- plot_rf + plot_apro + plot_DL +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = "A") & 
  theme(plot.tag = element_text(size = 10, face = "bold", family = "Helvetica"),
        legend.position="bottom",
        legend.spacing.x = unit(2, 'cm'),
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-25,0,-10,0))

ggsave("draft/figures/Fig3.pdf", fig3, width = 9.5, height = 2.8, device = cairo_pdf)
```

## Fig S1

can be found here: "results/ortho_bench/eggnog/accuracy.pdf" and can be reproduced
with any phylogenomics database annotated in uniprot following the instructions
in the main README.md

## Fig S2

```{r FigS2}
# Compare trees with RF distance
rf_df <- get_rf_df(filter(df_trees, model %in% models_draft))
rf_text <- group_by(rf_df, model.x, model.y) %>% 
  # filter(as.integer(model.x)<=as.integer(model.y)) %>% 
  summarise(value=mean(RF, na.rm=T))

plot_bs <- df_bs %>% 
  filter(model %in% c("LG", "GTR", "3Di", "AF", "LLM")) %>% 
  mutate(targets=target,
         model=factor(model, levels=models)) %>% 
  ggplot(aes(model, support, fill=targets)) +
  geom_boxplot(outlier.size = .1) +
  scale_fill_manual(values = palettes_method) +
  theme_classic() +
  labs(y="Boostrap", x="", fill="") +
  theme(legend.position = "bottom")

var_r2t <- reco %>% 
  filter(targets=="common", model %in% models_draft) %>% 
  ggplot(aes(variance_r2t, fct_reorder(model, variance_r2t))) +
  geom_violin(scale = "width", color="black", fill="grey40") +
  scale_x_continuous(limits = c(0,.5), expand = expansion(mult=0, add = c(0,0.05))) +
  labs(x="Variance Root-to-tip distance", y="") +
  # scale_fill_manual(values = palettes_model) +
  theme(legend.position = 'none',
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-10,0,0,0))

rf_plot <- rf_text %>% 
  ggplot(aes(model.x, model.y)) +
  geom_tile(aes(fill=value)) +
  geom_text(aes(label=round(value, 2)), color="white", size=3) +
  scale_fill_distiller(palette="YlGnBu", direction = -1, limits = c(0,1), 
                       name="Mean RF") + 
  labs(x="Mean RF Distance", y="") +
  theme_classic() +
  theme(legend.position = "none", 
        axis.line = element_blank(),
        axis.ticks = element_blank())

plot_struct_ml <- df_ml %>% 
  filter(model %in% c("GTR", "3Di", "AF", "LLM")) %>% 
  select(gene, targets, model, AIC, AICc, BIC) %>% 
  pivot_longer(!c(gene, targets, model)) %>% 
  group_by(gene, targets, name) %>% 
  mutate(rank = dense_rank(value)) %>% 
  filter(rank==1) %>% 
  ggplot(aes(model)) + 
  geom_bar(fill="grey40") + 
  scale_y_continuous(expand = expansion(mult = c(0,.02))) +
  facet_grid(~name) + 
  labs(x="", y="N. first in metric")

union_trees <- df_trees %>% 
  filter(target=="union", model %in% models_draft)
df_dist <- distance_to_seed(union_trees)

df_dist_summ <- df_dist %>% 
  left_join(select(df_filt, query, target, singleton), 
            by = c("query", "target")) %>%
  filter(query!=target, !is.na(singleton)) %>% 
  group_by(query) %>% 
  mutate(max_dist = max(nn_dist)) %>% 
  group_by(query, singleton, model) %>%
  summarise(median_dist=median(nn_dist/max_dist), n=n())

plot_dist <- df_dist_summ %>% 
    mutate(model=factor(model, levels=models)) %>% 
  ggplot(aes(model, median_dist,  fill=singleton)) + 
  geom_boxplot(
    outlier.color = NA,
    alpha = 1
  ) + 
  labs(y="Normalised internode distance", y="Method", fill="") +
  scale_y_continuous(expand = c(0,0.01), limits = c(0,1)) +
  theme(panel.border = element_rect(fill = NA)) + 
  scale_fill_manual(values = palette_singleton_p) + 
  theme(legend.position = "bottom")

first_row <- plot_bs + rf_plot + plot_struct_ml 
second_row <- plot_dist + var_r2t

fig_s2 <- (first_row / second_row) +
  plot_annotation(tag_levels = 'A') &
  theme(plot.tag = element_text(size = 10, face = "bold", family = "Helvetica"))

# fig_s2 <- ((plot_bs / rf_plot) + 
#              plot_layout(heights = c(2,1)) | 
#              (plot_dist / var_r2t) + 
#              plot_layout(heights = c(2,1))) +
#   plot_layout(widths = c(1, 2)) +
#   plot_annotation(tag_levels = 'A') & 
#   theme(plot.tag = element_text(size = 10, face = "bold", family = "Helvetica"))

ggsave("draft/figures/Fig_S2.pdf", fig_s2, width = 12, height = 8, device = cairo_pdf)

df_dist_summ %>% 
  group_by(singleton, model) %>% 
  summarise(median(median_dist)) %>% 
  filter(model=="LG")
```

## Fig S3

```{r, figS3}
bench_dir <- "draft/foldtree_benchmark/zenodo/OMA_data/Eukaryota/reconciliation/"
sptree_ft <- read.tree(paste0(bench_dir, "sptree.nwk"))

ts_files_ft <- list.files(paste0(bench_dir, "trees/tmp"), pattern = "*nwk", full.names = T)
ts_ft <- read.tree(text = sapply(ts_files_ft, readLines))
names(ts_ft) <- gsub(".nwk", "", basename(ts_files_ft))

n_tips_ft <- sapply(ts_ft, function(x) length(x$tip.label)) %>% 
  enframe() %>%
  separate(name, c("gene", "mat"), sep = "_") %>%
  rename("n_tips"="value")

TCS_ft <- read_delim(paste0(bench_dir,"TCS.txt"),
           delim = "\t", col_names = c("gene", "mat", "TCS")) %>% 
  mutate(mat = str_trim(mat))

apro_files_ft <- list.files(bench_dir, pattern = "^apro", full.names = T)
apro_trees_ft <- read.tree(text = sapply(apro_files_ft, readLines))
names(apro_trees_ft) <- gsub(".txt", "", sapply(str_split(basename(apro_files_ft), "_"), 
                            function(x) paste0(x[2])))
trees_df_ft <- get_apro_stats(apro_trees_ft)

disco_files_ft <- list.files(bench_dir, pattern = "^disco", full.names = T)
disco_rf_ft <- get_disco_rf(disco_files_ft, sptree_ft) %>% 
  mutate(mat = gsub(".txt", "", targets)) %>% 
  select(-targets, -alphabet, -model)

notung_files_ft <- list.files(bench_dir, pattern = "^notung", full.names = T)[2:5]

notung_ft <- read_delim(notung_files_ft, 
                   show_col_types = FALSE, 
                   col_names = c("gene", "dups", "losses"), 
           id = "mat") %>% 
  mutate(mat = gsub("notung_|.txt", "", basename(mat))) %>% 
  left_join(n_tips_ft) %>% 
  mutate(n_events = (dups+losses)/n_tips)

apro_plot_ft <- trees_df_ft %>%
  mutate(targets = gsub("iq", "iqtree", targets)) %>% 
  ggplot(aes(fct_rev(fct_reorder(targets, q1)), q1)) + 
  geom_boxplot(outlier.size = .1) + 
  labs(x="", y="Quartet support")

disco_plot_ft <- disco_rf_ft %>%
  mutate(mat = gsub("iq", "iqtree", mat)) %>%
  ggplot(aes(fct_reorder(mat, rf), rf)) + 
  geom_boxplot(outlier.size = .1) + 
  labs(y="Norm. RF to species tree", x="")

notung_plot_ft <- notung_ft %>%
  mutate(mat = gsub("iq", "iqtree", mat)) %>% 
  ggplot(aes(fct_reorder(mat, n_events), n_events)) + 
  geom_boxplot(outlier.size = .1) + 
  labs(y="(D+L)/# Tips", x="")

TCS_plot <- TCS_ft %>% 
  ggplot(aes(fct_rev(fct_reorder(mat, TCS)), TCS)) + 
  geom_boxplot(outlier.size = .1) + 
  scale_y_log10() + 
  labs(x="")

TCS_scatter <- TCS_ft %>% 
  filter(mat %in% c("iqtree", "fident")) %>% 
  pivot_wider(names_from = mat, values_from = TCS) %>% 
  mutate(is_better = iqtree > fident) %>% 
  add_count(is_better) %>% 
  mutate(`IQ-Tree Better` = paste0(is_better, ' (', n, ')')) %>% 
  # count(is_better)
  ggplot(aes(iqtree, fident, color=`IQ-Tree Better`)) + 
  geom_point() + 
  geom_abline(slope = 1) + 
  coord_fixed() + 
  labs(x="TCS - IQ-Tree", y="TCS - Fident") +
  scale_color_manual(values = c("blue", "red"))

left_side <- ((apro_plot_ft | disco_plot_ft) / (notung_plot_ft | TCS_plot))
fig_S3 <- (left_side | TCS_scatter) +
  plot_layout(widths = c(1,1)) +
  plot_annotation(tag_levels = 'A') &
  theme(plot.tag = element_text(size = 10, face = "bold", family = "Helvetica"))

ggsave("draft/figures/Fig_S3.pdf", fig_S3, width = 11, height = 5, device = cairo_pdf)
```

## Fig S4

```{r figs4}
plot_PDB <- threed_df %>%
  group_by(mnemo) %>% 
  filter(!is.na(mnemo)) %>% 
  summarise_all(~ (1-sum(is.na(.))/n())*100) %>% 
  left_join(select(tax, mnemo, Species)) %>% 
  ggplot(aes(PDB, fct_reorder(Species, PDB))) + 
  geom_text(aes(label=signif(PDB,2)), 
            size=2.5, nudge_x = 1, hjust = 0) + 
  geom_bar(stat="identity") + 
  scale_x_continuous(limits = c(0,100), expand = expansion(add = c(0,2))) +
  labs(y="Species", x="% Proteins with PDB entry") + 
  theme(axis.text.y = element_text(face = "italic"))

ggsave("draft/figures/Fig_S4.pdf", plot_PDB, width = 4, height = 3, device=cairo_pdf)
```

# Review plots

### Paralogy issue in phylome

```{r paralogy_phylome}
# find results/phylogeny/hsap_1kseeds/seeds/UP000005640/ -name "*.ids" -print0 | while IFS= read -r -d '' file; do  awk -v fname="$(basename "$file" | sed 's/.ids//' | sed 's/_/\t/')"  '{print $0 "\t" fname}' "$file"; done > test/paranoia/ids_by_set.txt
df_ids <- read_delim("draft/figures/ids_by_set.txt", 
                 col_names = c("target", "query", "set")) %>% 
  filter(query %in% aln_seeds)

overlap_df <- filter(df_ids, query!=target) %>% 
  group_by(query, set) %>% 
  summarise(targets=list(target)) %>% 
  # head(100) %>%
  cross_join(., .) %>% 
  filter(set.x==set.y, query.x!=query.y) %>% 
  rowwise() %>% 
  mutate(n =  length(union(targets.x, targets.y)), 
         overlap = length(intersect(targets.x, targets.y))/n) %>% 
  select(query.x, query.y, overlap, n, set.x) %>% 
  filter(query.x>query.y)

overlap_df %>% 
  mutate(is_not0 = overlap>0) %>% 
  count(set.x, is_not0) %>% 
  pivot_wider(names_from = is_not0, values_from = n) %>% 
  mutate(proportion = `TRUE`/(`FALSE`+`TRUE`)*100)
```

### Heteroscedascity of DL normalization

```{r heteroscedasticity}
lm <- lm(dups+losses ~ n_tips, data = reco)
reco$res <- lm$residuals

tmp_plot <- ggplot(reco, aes(y = res, x = n_tips, fill=target)) + 
  geom_point(pch=21) +
  facet_wrap(~target) + 
  geom_abline(slope = 0) + 
  scale_fill_manual(values = palettes_method) + 
  labs(x="Num. of Tips", 
       y="Residuals of: D+L ~ Num. of Tips") + 
  theme(legend.position = "none")

ggsave("draft/figures/review_hetero_TMP.png", tmp_plot,
       width = 4, height = 3)
```

### CXX models

```{r cxx_models}
models_cxx <- c("LG", "C10", "C60", "FM", "FMG")
times_df <- read_delim("test/flashforward/times.txt") %>% 
  mutate(max_rss_gb = max_rss/1024) %>% 
  separate(file, c("seed", "gene", "targets", "alphabet", "model"), sep = "_") %>% 
  select(gene, targets, model, s, `h:m:s`, max_rss_gb)

seeds_all <- filter(df_trees, model=="C60") %>% pull(gene)

plot_time_cxx <- times_df %>% 
  # filter() %>% 
  ggplot(aes(fct_reorder(model, s), s/3600)) + 
  geom_boxplot(outlier.size = .1) +
  labs(x="", y="# hours")

plot_mem_cxx <- times_df %>% 
  ggplot(aes(fct_reorder(model, max_rss_gb), max_rss_gb)) + 
  geom_boxplot(outlier.size = .1) +
  labs(x="", y="Memory usage (GB)")

DL_cxx <- reco %>% 
  left_join(df_model) %>% 
  filter(model %in% models_cxx, targets=="common") %>% 
  ggplot(aes(model, n_events, fill=targets)) + 
  labs(y="(D+L)/# Tips", x="",fill="Target sets") +
  geom_boxplot(outlier.size = .1) + 
  scale_fill_manual(values = palettes_method)

RF_cxx <- disco_rf %>% 
  mutate(model=factor(model, levels=models)) %>% 
  filter(nsps>length(sptree$tip.label)/2, model %in% models_cxx, targets=="common") %>%
  ggplot(aes(model, rf, fill=targets)) + 
  geom_boxplot(outlier.size = .1) +
  labs(x="", y="Norm. RF", fill="Target sets") +
  scale_fill_manual(values = palettes_method)

apro_CXX <- trees_df %>% 
  filter(model %in% models_cxx, targets=="common") %>% 
  ggplot(aes(model, q1, fill=targets)) + 
  labs(x="", y="Quartet support", fill="Target sets") + 
  geom_boxplot(outlier.size = .1) +
  scale_fill_manual(values = palettes_method)

fig_CXX <- (RF_cxx + apro_CXX + DL_cxx + plot_mem_cxx + plot_time_cxx)
fig_CXX <-  fig_CXX +
  plot_layout(guides = "collect", nrow = 1, 
              widths = c(5,5,5,3,3)) +
  plot_annotation(tag_levels = 'A') &
  theme(plot.tag = element_text(size = 10, face = "bold", family = "Helvetica"))

ggsave("draft/figures/review_CXX.pdf", fig_CXX, width = 9.5, height = 2.8, device = cairo_pdf)
```

### Query coverage and BlastP singletons

```{r qcov_blastpsingletons}
pfam_jaccard <- df %>% 
  # head(100000) %>% 
  select(query, target, qcov_Blast, qcov_Foldseek, singleton) %>% 
  left_join(select(ortho_dbs, target, Pfam), by=c("query"="target")) %>% 
  left_join(select(ortho_dbs, target, Pfam), by=c("target")) %>% 
  rename("query_set"=paste0("Pfam", ".x"), "target_set"=paste0("Pfam", ".y")) %>% 
  # If any of query or target has na no we dont count
  filter(!is.na(query_set), !is.na(target_set)) %>% 
  mutate(
    jaccard = map2_dbl(strsplit(query_set, ";"), strsplit(target_set, ";"),
                       ~ length(intersect(.x, .y)) / length(union(.x, .y)))
  )

plot_pfam_qcov <- pfam_jaccard %>% 
  mutate(qcov = rowMeans(select(., qcov_Blast, qcov_Foldseek), na.rm = TRUE)) %>% 
  mutate(disc=cut(qcov, breaks = seq(0,100,25))) %>% 
  ggplot(aes(jaccard, disc, color=singleton)) + 
  geom_boxplot(outlier.size = .1) +
  labs(y="Q. cov. interval", x="Pfam JI") +
  scale_color_manual(values=palette_singleton_p)

ggsave("draft/figures/review_qcov.pdf", plot_pfam_qcov, width = 5, height = 6, device = cairo_pdf)

# query_pfam <- pfam_jaccard %>% 
#   # filter(query=="AF-P06681-F1") %>% 
#   group_by(query) %>% 
#   summarise(#all_domains = paste0(target_set, collapse = ""),
#             q_domains = paste0(unique(query_set), collapse = "")) %>% 
#   separate(q_domains, letters, sep = ";")  %>% 
#   pivot_longer(cols = !c(query), values_drop_na = T)  %>% 
#   filter(value!="") %>% 
#   select(-name)
# 
# top_pfam <- pfam_jaccard %>% 
#   # filter(query=="AF-P06681-F1") %>%
#   select(query, target_set)  %>% 
#   left_join(query_pfam, relationship = "many-to-many") %>% 
#   mutate(is_in = str_detect(target_set, value)) %>%
#   group_by(query, value) %>% 
#   summarise(n = sum(is_in)) %>% 
#   slice_max(n, with_ties = FALSE)
# 
# prop_topdomain <- pfam_jaccard %>% 
#   # filter(query=="AF-A0A024RCN7-F1") %>% 
#   select(query, target_set, singleton) %>% 
#   left_join(top_pfam) %>% 
#   mutate(is_in = str_detect(target_set, value)) %>% 
#   group_by(query, singleton) %>% 
#   mutate(n_group = n()) %>% 
#   summarise(n = sum(is_in)/mean(n_group))
# 
# prop_topdomain %>% 
#   ggplot(aes(n, singleton)) + 
#   geom_boxplot() + 
#   labs(x="% with top domain")
```